---
title: "Cunningham - TidyVerse ggplot2"
author: "Brandon Cunningham"
date: "`r Sys.Date()`"
output: html_document
---

# Intro
I am using data from the Kaggle dataset [Fuel Economy](https://www.kaggle.com/datasets/sahirmaharajj/fuel-economy)
which I have uploaded to [github](https://github.com/btc2628/DATA607/blob/main/fuel_data/fuel.csv) for easy access to the data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(knitr)
library(dplyr)
library(readxl)
library(httr)
library(tidyverse)
library(ggplot2)
```

In this vignette I will be using data on fuel economy to show off features of ggplot2 and dplyr.

```{r}
url <- 'https://raw.githubusercontent.com/btc2628/DATA607/main/fuel_data/fuel.csv'
fuel_raw <- readr::read_csv(url)
fuel_narrowed <- fuel_raw[, c('make', 'model', 'year', 'class', 'drive', 'transmission', 'engine_cylinders', 'engine_displacement', 'turbocharger', 'fuel_type', 'combined_mpg_ft1')]
```


Now we have the data loaded, however while looking at the raw data I noticed there were a lot of different transmission types, we can use dplyr to figure out just how many there are and get a good view of them using dplyr::select() to choose the column we would like to inspect, and dplyr::distinct() to get all distinct values from that column.
```{r}
transmission_sample <- fuel_narrowed %>%
  dplyr::select(transmission) %>%
  dplyr::distinct()
transmission_sample
```
As we can see, there are 47 distinct values in the transmission column including NA. However what we might want to do for easier analysis is narrow this down to their major categories by using the mutate function in dplyr to grab only the first word from the column which in this case would be the transmission type, and filter function to remove any NA values.

```{r}
fuel_narrowed <- fuel_narrowed %>%
  dplyr::filter(!is.na(transmission)) %>%
  dplyr::mutate(transmission = str_extract(transmission, "^[^\\s\\(]+"))

transmission_new <- fuel_narrowed %>%
  dplyr::select(transmission) %>%
  dplyr::distinct()
transmission_new
```
After mutating with dplyr we are able to narrow down the number of transmission types to 3 which makes the data a lot more usable for analysis.  

And now we can use ggplot2 to create a graph for each of the 3 types of transmission we found plotting the mean mpg of each for each year we have data for. 
```{r}
ggplot(fuel_narrowed, aes(x=year, y=combined_mpg_ft1, group=transmission, fill=transmission)) + 
   geom_bar(stat="summary", fun="mean", show.legend = FALSE) +
  facet_wrap(~transmission, scales = 'free_x') + 
  labs(x='Year', y='Combined MPG') 
```

# Richie Extension

From the graph above, it looks like there isn't a great difference between Automatic/Auto and Manual in the modern day. There are even observations where auto/automatic transmissions are outperforming manual ones!

I wonder if there are other categorical variables where we would see a different trend. Using the same method to create a similar graph with vehicle type:

```{r richie}
str(fuel_narrowed)

ggplot(fuel_narrowed, aes(x=year, y=combined_mpg_ft1, group=class, fill=class)) + 
   geom_bar(stat="summary", fun="mean", show.legend = FALSE) +
  facet_wrap(~class, scales = 'free_x') + 
  labs(x='Year', y='Combined MPG') 
```
Because there are so many categories for `class` it's really difficult to compare across classes. But we can see that the broad trend is that newer cars generally became more efficient. With few exceptions, but notably Vans in general have become noticeably less efficient.

Plotting this another way:

```{r richie}
avg_mpg_hist_class <- fuel_narrowed |>
  group_by(
    year,
    class
  ) |>
  summarise(
    avg_mpg = mean(combined_mpg_ft1)
  )

ggplot(
  avg_mpg_hist_class,
  aes(
    x = year,
    y = avg_mpg,
    color = class
  )
) +
  geom_line() +
  labs(
    title = "Annual average MPG by class and year",
    x = "Year",
    y = "Average MPG"
  ) +
  theme_minimal()
```

This graph is crazily noisy, and it seems like it's due to the lack of entries for some vehicle types across time.

Lastly, let's pull out which class is the best for the most recent year of data:

```{r richie}
max_year <- max(avg_mpg_hist_class$year)

latest_year <- avg_mpg_hist_class |>
  filter(
    year == max_year
  ) |>
  arrange(
    desc(
      avg_mpg
    )
  )

latest_year
```

Pretty expectedly, this would be the Small Station wagons and Midsize cars with the gap between the two being fairly large.