---
title: "DATA 607 TidyVerse Create"
author: "Biyag Dukuray"
output: html_document
date: "2024-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 607 TidyVerse Create

In this assignment, you’ll practice collaborating around a code project with GitHub. You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository: <https://github.com/acatlin/SPRING2024TIDYVERSE>

FiveThirtyEight.com datasets.

Kaggle datasets.

Your task here is to Create an Example. Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)

Later, you'll be asked to extend an existing vignette. Using one of your classmate’s examples (as created above), you'll then extend his or her example with additional annotated code. (15 points)

You should clone the provided repository. Once you have code to submit, you should make a pull request on the shared repository. You should also update the README.md file with your example.

After you’ve created your vignette, please submit your GitHub handle name in the submission link provided below. This will let your instructor know that your work is ready to be peer-graded.

You should complete your submission on the schedule stated in the course syllabus.

**The dataset that I decided to use to complete this assignment and demonstrate the uses and capabilities of the TidyVerse package is the "Insurance claims and policy data" from kaggle. This dataset is a a comprehensive collection designed to facilitate predictive analytics and risk assessment within the insurance industry. My theory is that people who are experiencing stressful life situations are more likely to have higher claims history thus resulting in a higher risk of policy payout. I believe indicuals who are experienceing either a seperation, divorce or widowed or more likely to have a higher claims history in comparison to people who are married or single. We will begin by loading the libraries needed to demonstrate the capabilities of the Tidyverse package as well as loading the insurance data into a dataframe.**

**Link to dataset:** <https://www.kaggle.com/datasets/ravalsmit/insurance-claims-and-policy-data>

```{r}
library(tidyverse)
library(readr)
```

```{r}
kaggleurl <- "https://raw.githubusercontent.com/Zcash95/DATA607-files/main/data_synthetic.csv"

insurance_data <- read_csv(kaggleurl)
```

**We will now take a look at the dataset to see what attributes we can use to assess and gauge our theory.**

```{r}
glimpse(insurance_data)
```

**I will use ggplot to help create a plot to visualize the relationship statuses of the insured.**

```{r}
marital_plot <- insurance_data %>% count(`Marital Status`) %>%
ggplot(aes(x = `Marital Status`, y = n)) + geom_bar(stat = "identity", fill = "skyblue") +
labs(title = "Distribution of Marital Status", x = "Marital Status", y = "Count") + theme_minimal()
print(marital_plot)
```

**We can see that there is about an equal amount of people who are married or divorced but there are less people who are separated, single or widowed. We will now analyze whether there is a correlation between a persons relationship status and the amount of claims by persons in this group.**

```{r}
correlation <- insurance_data %>%
mutate(`Marital Status` = factor(`Marital Status`)) %>%
group_by(`Marital Status`) %>%
summarise(mean_claim = mean(`Claim History`)) %>%
arrange(desc(mean_claim))


print(correlation)
```

**We can see the correlation shows that single and married people have less of a claims history in comparison to people who are separated, widowed or divorced. We will now use ggplot to visualize this correlation as well ordering the bars in the same order to demonstrate which relationship status is more likely to have a higher claims history thus being a higher risk for an insurance company.**

```{r}
correlation_plot <- correlation %>%
ggplot(aes(x = reorder(`Marital Status`, -mean_claim), y = mean_claim, fill = `Marital Status`)) +
geom_bar(stat = "identity") +
labs(title = "Mean Claim Amount by Marital Status", x = "Marital Status", y = "Mean Claim Amount") + 
theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(correlation_plot)
```

**We can also show the distribution of number of claims and break down each bar in the graph by the marital status which made that number of claims. Through this we can see if there are patterns in any specific number of claims made by certain martial status groups, however there does not appear to be much of a relationship between the marital status and any specific number of claims**
```{r}
ggplot(insurance_data, aes(x=`Claim History`, fill=`Marital Status`)) + 
   geom_bar() +
  labs(x='Claim History')
```

**Another thing we might want to look at is if a person's coverage amount is correlated with their claim history. to do this we have a few options, we could create a bar graph with a persons claim history on the x-axis and average coverage on the y axis, but as there have already been quite a few bar graph examples we will use another option. Another way of showing this would be to use a scatter plot with the x-axis representing the number of claims they've made, and y axis representing their coverage amount.**
```{r}
ggplot(insurance_data, aes(x=`Claim History`, y=`Coverage Amount`)) + 
   geom_point() +
  labs(x='Claim History', y='Coverage Amount')
```
**Here we have a problem, there are too many data points and too few categories that what we end up with just looks like a giant straight line in each category. To fix this there are a couple options, the first is we can add a transparency to each point so that where there are more data points in one spot it will appear more visible**
```{r}
ggplot(insurance_data, aes(x=`Claim History`, y=`Coverage Amount`)) + 
   geom_point(alpha=0.01) +
  labs(x='Claim History', y='Coverage Amount')
```
**Adding transparency helps, but it still has a problem of being a perfectly straight thin line which makes it hard to get a graph of the number of points in each coverage level. What we can do to help alleviate this is use geom_jitter instead of geom_point to make it so that the points are distributed around their claim history amount instead of perfectly on it to turn the straight line into more of a bar.**
```{r}
ggplot(insurance_data, aes(x=`Claim History`, y=`Coverage Amount`)) + 
   geom_jitter(alpha=.1) +
  labs(x='Claim History', y='Coverage Amount')
```
**After adding these visual adjustments, we can much more clearly see the different densities of points in each claim group and we can even see stripes of points at popular coverage amounts.**

**Conclusion: GGplot2 gave us the ability to display this analysis of the data and the correlation between a persons relationship status and their insurance claims history. This confirmed my theory that people who are in stable situations such as being married or single will have less insurance claims in comparison to people who may be in stressful situations such as being separated, widowed or divorced. Though this is the case, this should not be the only factor analyzed by an insurance company to gauge the risk of a consumer.**
