---
title: "TIdyverse Assignment"
author: "Jonathan Cruz"
date: "2024-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Task

I would like to assess the text-to-sentence ratio in each text for a Large Language Model (LLM).
Following that, I aim to analyze the overall sentiment expressed within each text.

```{r include=FALSE}
library(tidyverse)
library(knitr)
library(tidytext)
```

# Data 


## Source 
https://www.kaggle.com/datasets/carlmcbrideellis/llm-7-prompt-training-dataset

## Download Link 
https://www.kaggle.com/datasets/carlmcbrideellis/llm-7-prompt-training-dataset/download?datasetVersionNumber=4

## Files 
train_essays_7_prompts.csv
train_essays_7_prompts_v22.csv


```{r load}

dt1  <- read_csv("train_essays_7_prompts.csv")
dt2  <- read_csv("train_essays_7_prompts_v2.csv")
```



You can consolidate rows from multiple dataframes into one dataframe using the [dplyr]{.underline} package in R. This allows you to merge dataframes based on their similarities.

```{r bind_rows}
dt <- bind_rows(dt1, dt2)
```


With the [stringr]{.underline} package in R, you can analyze text in each row of a dataframe and match them with expressions. This enables you to perform various text processing tasks, such as pattern matching and extracting substrings.


```{r}
dt <- dt |>
          mutate(total_words = str_count(text,'\\w+ ')+1) |>
          mutate(total_sentences = str_count(text, "\\.|\\?" )) |>
          mutate(ratio = total_words / total_sentences) |>
          mutate(id = row_number())
          
```


I will utilize [tidytext]{.underline}, which is built using tidyverse tools, to extract sentiment values using [dplyr]{.underline} in conjunction with [tidytext]{.underline}.


```{r dyplyr_stringr, echo=FALSE}
sentiments <- get_sentiments("afinn")
sent_dt <- dt |>
           select(text, id) |>
           mutate( text = str_remove_all(text, "\\.|\\?|\"|'|,|:")) |>
           unnest_tokens(word, text) |>
           inner_join(sentiments) |> 
           group_by(id) |>
           summarise(net = sum(value))
```
more [dplyr]{.underline} functions

```{r inner_join}
text_sent <- inner_join(dt, sent_dt)
```



We can enhance our data interpretation by employing [ggplot]{.underline} to create insightful visualizations.

```{r ggplot}
ggplot(text_sent, aes(x=total_words)) + 
    geom_histogram(aes(y=after_stat(density)),     
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
```
