---
title: "TIDYVERSE Assignment"
author: "angel"
extended author: "Dhanya nair"
date: "2024-04-07"
Extended date: "2024-04-21"
output: html_document
---

## Task: to create a vignette that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

# The data set that I will be using is the current homes for sale in NY
#https://www.kaggle.com/datasets/nelgiriyewithana/new-york-housing-market

#I will be demonstrating how to use  package with readr, dplyr,ggplot2 
```{r}
library(tidyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

```


#read the csv for data on current house for sales in NY
```{r}
df <- read.csv("https://raw.githubusercontent.com/Angelogallardo05/Data606-proposal/main/NY-House-Dataset.csv")
```

```{r}
head(df,3)
```

# dplyr

#Remove NA 
```{r}
df <- df %>%
  na.omit(df)%>%
  mutate(PRICE = as.numeric(PRICE))
```

#mutate to combine Queens County with Queens
```{r}
df <- df %>%
  mutate(SUBLOCALITY = case_when(
    SUBLOCALITY == "Queens" ~ "Queens County",
    TRUE ~ SUBLOCALITY  
  ))
```


#filter the data to remove outliers, and the type of home for sale to be a house in Queens
```{r}
filtered_df <- df %>%
  filter(PROPERTYSQFT != 2184.207862)%>%
  filter(TYPE == "House for sale", SUBLOCALITY == "Queens County")

# Check the filtered dataframe

```


#Summerize the average price, beds, baths and sqft. Average house for sale price in Queens is $1.1MM with about 4 beds and 2.5 baths and 1,957 sqft
```{r}
summary_stats <- filtered_df %>%
  summarise(
    count = n(),                             
    avg_price = mean(PRICE, na.rm = TRUE),   
    median_price = median(PRICE, na.rm = TRUE),  
    min_price = min(PRICE, na.rm = TRUE),     
    max_price = max(PRICE, na.rm = TRUE),     
    avg_beds = mean(BEDS, na.rm = TRUE),      
    avg_bath = mean(BATH, na.rm = TRUE),      
    avg_sqft = mean(PROPERTYSQFT, na.rm = TRUE)  
  )

# View the summary statistics
print(summary_stats)
```


# ggplot2


#use ggplot to visualize the correlation between Price and sqft, price and beds, Price and baths in queens. Seems like sqfootage and Price have a linear correlation
```{r}
ggplot(filtered_df, aes(x = PROPERTYSQFT, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Property Square Footage", y = "Price", title = "Price vs. Property Square Footage") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()

ggplot(filtered_df, aes(x = BEDS, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Beds", y = "Price", title = "Price vs. Beds") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()

ggplot(filtered_df, aes(x = BATH, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Baths", y = "Price", title = "Price vs. Baths") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()
```


## Extension of the code by Dhanya Nair
## 1. Tidying the data
### In the BROKERTITLE variable, there are 35 records which  has " No Broker" and few records which are " Built by Toll Brothers".

### I want to keep only brokers name in this column.

### Using separate(), drop_na() and filter()
```{r}
# to remove the text "Brokered by" from the BROKERTITLE variable
broker_df <- df |>
separate(BROKERTITLE, into = c("prefix", "BROKER"), sep = "\\sby\\s", convert = TRUE)

# to remove NA , this will filter out the No Broker records.
# Remove " Built by Toll brothers" as these records will have "Built" in the prefix column
# hence fitlering by "Brokered"
broker_df <- broker_df |>
  drop_na(BROKER) |>
  filter(prefix == "Brokered")



```


2. using round() function

```{r}

broker_df$BATH <- round(broker_df$BATH,0)
broker_df$PROPERTYSQFT <- round(broker_df$PROPERTYSQFT,0)


```


## 3. which brokerage sold the most homes in New York ?
### using table() , sort() and head()
```{r}

# Use table() function to summarize and count the values
broker_stat <- table(broker_df$BROKER)

# Sort the summary table by counts
broker_stat <- sort(broker_stat, decreasing = TRUE)

# Display the top 10 values
top_10 <- head(broker_stat, 10)

# Print the top 10 values
print(top_10)

```

## 4. Visualize the data for PRICE by HOUSE TYPE for over 6000 SQFT

### using box_plot()  and facet_wrap()
```{r}

broker_df_high<- broker_df |>
  filter(PROPERTYSQFT > 6000)
  
broker_df_high %>%
  drop_na(PRICE)%>%
   group_by(TYPE) %>%
  ggplot(aes(x = PRICE, group = TYPE, color = TYPE)) +
  geom_boxplot(outlier.shape = NA) +
    labs(title  = "HOUSE SOLD PRICE (USD) with over 6000 SQFT") +
    theme(axis.title = element_text(face = "bold"), 
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          strip.text.x = element_text(size = 8)) +  
  facet_wrap(~ TYPE) +
  
    scale_x_continuous(breaks = c(0, 5e5, 1e6, 1e7), 
                     limits = c(4e5, 1e8),
                     labels = scales::scientific)  +
  theme_minimal()


```





## Extra:

### 4. display the houses sold in Queens, NY in a map
```{r}
library(maps)
library(ggmap)
library(tigris) # for accessing geographic data


# Get boundary data for Queens County, New York
queens_county <- tigris::counties(state = "NY", cb = TRUE) %>%
  filter(NAME == "Queens")

# Plot the map of Queens County, New York, with houses
ggplot() +
  geom_sf(data = queens_county, fill = "white", color = "black") +
  geom_point(data = filtered_df,
             aes(x = LONGITUDE, y = LATITUDE),
             color = "red", size = 3) +
  labs(x = "Longitude", y = "Latitude", title = "Map of Queens County, NY with Houses sold") +
  theme_minimal()



```

## Conclusion:
- I have extended Angel's assignment . 
- I worked on tidying the data using Using separate(), drop_na() and filter(),round() . 
- I created summaries using  table() , sort() and head() .
- I visualized the data using box_plot()  and facet_wrap().
- Additionally i displayed the data using a package not in tidyverse using maps.

